---
title: "Lendingclub"
author: "elsa"
date: "8 septembre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
library(tidyverse)
library(data.table)
library(stringr)
library(DescTools)
library(lubridate)
```

revol_bal: amount used and not repaid
revol_util:the amount of credit the borrower is using relative to all available revolving credit.
total_acc: The total number of credit lines currently in the borrower's credit file
initial_list_status: two possible settings –“F” for fractional, “W” for whole.
out_prncp:	Remaining outstanding principal for total amount funded
out_prncp_inv	Remaining outstanding principal for portion of total amount funded by investors
total_pymnt	Payments received to date for total amount funded
total_pymnt_inv	Payments received to date for portion of total amount funded by investors
total_rec_int	Interest received to date
total_rec_late_fee	Late fees received to date
total_rec_prncp	Principal received to date
recoveries	post charge off gross recovery
collection_recovery_fee	post charge off collection fee
last_pymnt_d	Last month payment was received   Date
next_pymnt_d	next month payment was received  Date
last_credit_pull_d	The most recent month LC pulled credit for this loan: checked or decreased the score? Date
collections_12_mths_ex_med	Number of collections in 12 months excluding medical collections
recouvrement de dettes
mths_since_last_major_derog	Months since most recent 90-day or worse rating Month
policy_code	"publicly available policy_code=1 new products not publicly available policy_code=2"
application_type individual application or a joint application with two co-borrowers if individual alors tous ce qui est joint est vide
annual_inc_joint	The combined self-reported annual income provided by the co-borrowers during registration            addition
dti_joint	A ratio calculated using the co-borrowers' total monthly payments on the total debt obligations, excluding mortgages and the requested LC loan, divided by the co-borrowers' combined self-reported monthly income           addition
verification_status_joint: 
acc_now_delinq	The number of accounts on which the borrower is now delinquent.
tot_coll_amt	Total collection amounts ever owed
tot_cur_bal	Total current balance of all accounts
open_acc_6m	Number of open trades in last 6 months
open_il_6m	Number of currently active installment trades
open_il_12m	Number of installment accounts opened in past 12 months
open_il_24m	Number of installment accounts opened in past 24 months
open_rv_12m	Number of revolving trades opened in past 12 months
open_rv_24m	Number of revolving trades opened in past 24 months
mths_since_rcnt_il	Months since most recent installment accounts opened
total_rev_hi_lim 	Total revolving high credit/credit limit



```{r}
data_lending <- read_csv("../inputs/loan.csv", 
  col_types = cols(
  id = col_character(),
  member_id = col_character(),
  loan_amnt = col_double(),
  funded_amnt = col_double(),
  funded_amnt_inv = col_double(),
  term = col_character(),
  int_rate = col_double(),
  installment = col_double(),
  grade = col_character(),
  sub_grade = col_character(),
  emp_title = col_character(),
  emp_length = col_character(),
  home_ownership = col_character(),
  annual_inc = col_double(),
  verification_status = col_character(),
  issue_d = col_character(),
  loan_status = col_character(),
  pymnt_plan = col_character(),
  url = col_character(),
  desc = col_character(),
  purpose = col_character(),
  title = col_character(),
  zip_code = col_character(),
  addr_state = col_character(),
  dti = col_double(),
  delinq_2yrs = col_double(),
  earliest_cr_line = col_character(),
  inq_last_6mths = col_double(),
  mths_since_last_delinq = col_double(),
  mths_since_last_record = col_double(),
  open_acc = col_double(),
  pub_rec = col_double(),
  revol_bal = col_double(),
  revol_util = col_double(),
  total_acc = col_double(),
  initial_list_status = col_character(),
  out_prncp = col_double(),
  out_prncp_inv = col_double(),
  total_pymnt = col_double(),
  total_pymnt_inv = col_double(),
  total_rec_prncp = col_double(),
  total_rec_int = col_double(),
  total_rec_late_fee = col_double(),
  recoveries = col_double(),
  collection_recovery_fee = col_double(),
  last_pymnt_d = col_character(),
  last_pymnt_amnt = col_double(),
  next_pymnt_d = col_character(),
  last_credit_pull_d = col_character(),
  collections_12_mths_ex_med = col_double(),
  mths_since_last_major_derog = col_character(),
  policy_code = col_double(),
  application_type = col_character(),
  annual_inc_joint = col_character(),
  dti_joint = col_character(),
  verification_status_joint = col_character(),
  acc_now_delinq = col_double(),
  tot_coll_amt = col_character(),
  tot_cur_bal = col_character(),
  open_acc_6m = col_character(),
  open_il_6m = col_character(),
  open_il_12m = col_character(),
  open_il_24m = col_character(),
  mths_since_rcnt_il = col_character(),
  total_bal_il = col_character(),
  il_util = col_character(),
  open_rv_12m = col_character(),
  open_rv_24m = col_character(),
  max_bal_bc = col_character(),
  all_util = col_character(),
  total_rev_hi_lim = col_double(),
  inq_fi = col_character(),
  total_cu_tl = col_character(),
  inq_last_12m = col_character()
))
```

# Exploratory Data Analysis

```{r}
summary(data_lending)
```

## Missing values

```{r}
missing_table <-  data_lending %>% 
  map_dbl(~ round(sum(is.na(.x))/length(.x), digits=2)) %>%
  data_frame(a=names(.), b=.) %>% 
  DT::datatable()

delete_missing <- missing_table$x$data %>%  filter(b>0.7) %>% select(a) %>% unlist()
data_lending_deleted_missing <- data_lending[ , -which(names(data_lending) %in% delete_missing)]

missing_table

```

## Feature Cleaning

```{r}
#extract numbers from term
data_clean <- data_lending_deleted_missing %>% 
  mutate(term= as.numeric(str_extract_all
                          (term, "[[:digit:]]+")))

#extract numbers from employment length
data_clean1 <- data_clean %>% 
  mutate(emp_length= as.numeric(str_extract_all
                                (emp_length, "[[:digit:]]+")))

#Binary Verification Status
data_clean2 <- data_clean1 %>% 
  mutate(verification_status= ifelse
         (verification_status=="Not Verified", 0, 1)) 

#Issue Date into date format

months_dict <- c("01", "02", "03", "4", "5", "6", "7", "8", "9", "10", "11", "12")
names(months_dict) <- c("Jan", "Feb", "Mar", "Apr", "May",
                        "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

data_clean3 <- data_clean2 %>% 
  separate(issue_d, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(issue_d, d_d, good_m, y_d, sep="-") %>% 
  mutate(issue_d= dmy(issue_d)) %>% 
  select(-m_d)

#Earliest Credit Line into date format

data_clean4 <- data_clean3 %>% 
  separate(earliest_cr_line, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(earliest_cr_line, d_d, good_m, y_d, sep="-") %>% 
  mutate(earliest_cr_line= dmy(earliest_cr_line)) %>% 
  select(-m_d)

# Last payment into date format
  
data_clean5 <- data_clean4 %>% 
  separate(last_pymnt_d, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(last_pymnt_d, d_d, good_m, y_d, sep="-") %>% 
  mutate(last_pymnt_d= dmy(last_pymnt_d)) %>% 
  select(-m_d)

# Last Credit Pull into date format

data_clean6 <- data_clean5 %>% 
  separate(last_credit_pull_d, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(last_credit_pull_d, d_d, good_m, y_d, sep="-") %>% 
  mutate(last_credit_pull_d= dmy(last_credit_pull_d)) %>% 
  select(-m_d)
  
```


## Verification of mathematical formulas and Calculation of actual benefits for the investors

```{r}
# Formulas to calclulate monthly payment due 
data_clean6 %>% 
  mutate(monthly_pymt_calculated= (loan_amnt*(int_rate*0.01/12))/(1 - (1 +int_rate*0.01/12) ^(-term))) %>%      mutate(diff_pmt=monthly_pymt_calculated - installment) %>% 
  select(installment, monthly_pymt_calculated, diff_pmt)

# Real total payment and benefits for the investors

data_benef <- data_clean6 %>%
  mutate(real_pymt = installment*term) %>% 
  mutate(total_benef = real_pymt - loan_amnt)

# Investor theoretical ROI

data_ROI <- data_benef %>% 
  mutate(roi = (total_benef/loan_amnt))


```


# Graphs

```{r}
Desc(data_clean$loan_amnt, main = "Loan amount distribution", plotit = TRUE)
```

```{r}
Desc(data_clean$loan_status, plotit = TRUE)
```

## Theoretical Roi compared to Grades

```{r}
ggplot(data_ROI, aes(x= as.factor(loan_status), y=roi)) +
  geom_boxplot(aes(fill=as.factor(loan_status))) +
  theme(legend.position= "right")

```



## Theoretical Roi compared to actual status of loan

```{r}
ggplot(data_ROI, aes(x= as.factor(grade), y=roi)) +
  geom_boxplot(aes(fill=as.factor(grade))) +
  theme(legend.position= "right")

```



Quel proportion de mec voit son pret financer? 

Quel son les prêts qui font défaults

Rentabilité théorique vs réelle