---
title: "Lendingclub"
author: "elsa"
date: "8 septembre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
library(tidyverse)
library(data.table)
```
revol_bal: amount used and not repaid
revol_util:the amount of credit the borrower is using relative to all available revolving credit.
total_acc: The total number of credit lines currently in the borrower's credit file
initial_list_status: two possible settings –“F” for fractional, “W” for whole.
out_prncp:	Remaining outstanding principal for total amount funded
out_prncp_inv	Remaining outstanding principal for portion of total amount funded by investors
total_pymnt	Payments received to date for total amount funded
total_pymnt_inv	Payments received to date for portion of total amount funded by investors
total_rec_int	Interest received to date
total_rec_late_fee	Late fees received to date
total_rec_prncp	Principal received to date
recoveries	post charge off gross recovery
collection_recovery_fee	post charge off collection fee
last_pymnt_d	Last month payment was received   Date
next_pymnt_d	next month payment was received  Date
last_credit_pull_d	The most recent month LC pulled credit for this loan: checked or decreased the score? Date
collections_12_mths_ex_med	Number of collections in 12 months excluding medical collections
recouvrement de dettes
mths_since_last_major_derog	Months since most recent 90-day or worse rating Month
policy_code	"publicly available policy_code=1 new products not publicly available policy_code=2"
application_type individual application or a joint application with two co-borrowers if individual alors tous ce qui est joint est vide
annual_inc_joint	The combined self-reported annual income provided by the co-borrowers during registration            addition
dti_joint	A ratio calculated using the co-borrowers' total monthly payments on the total debt obligations, excluding mortgages and the requested LC loan, divided by the co-borrowers' combined self-reported monthly income           addition
verification_status_joint: 
acc_now_delinq	The number of accounts on which the borrower is now delinquent.
tot_coll_amt	Total collection amounts ever owed
tot_cur_bal	Total current balance of all accounts
open_acc_6m	Number of open trades in last 6 months
open_il_6m	Number of currently active installment trades
open_il_12m	Number of installment accounts opened in past 12 months
open_il_24m	Number of installment accounts opened in past 24 months
open_rv_12m	Number of revolving trades opened in past 12 months
open_rv_24m	Number of revolving trades opened in past 24 months
mths_since_rcnt_il	Months since most recent installment accounts opened
total_rev_hi_lim 	Total revolving high credit/credit limit



```{r}
data_lending <- read_csv("../input/loan.csv", 
  col_types = cols(
  id = col_character(),
  member_id = col_character(),
  loan_amnt = col_double(),
  funded_amnt = col_double(),
  funded_amnt_inv = col_double(),
  term = col_character(),
  int_rate = col_double(),
  installment = col_double(),
  grade = col_character(),
  sub_grade = col_character(),
  emp_title = col_character(),
  emp_length = col_character(),
  home_ownership = col_character(),
  annual_inc = col_double(),
  verification_status = col_character(),
  issue_d = col_character(),
  loan_status = col_character(),
  pymnt_plan = col_character(),
  url = col_character(),
  desc = col_character(),
  purpose = col_character(),
  title = col_character(),
  zip_code = col_character(),
  addr_state = col_character(),
  dti = col_double(),
  delinq_2yrs = col_double(),
  earliest_cr_line = col_character(),
  inq_last_6mths = col_double(),
  mths_since_last_delinq = col_double(),
  mths_since_last_record = col_double(),
  open_acc = col_double(),
  pub_rec = col_double(),
  revol_bal = col_double(),
  revol_util = col_double(),
  total_acc = col_double(),
  initial_list_status = col_character(),
  out_prncp = col_double(),
  out_prncp_inv = col_double(),
  total_pymnt = col_double(),
  total_pymnt_inv = col_double(),
  total_rec_prncp = col_double(),
  total_rec_int = col_double(),
  total_rec_late_fee = col_double(),
  recoveries = col_double(),
  collection_recovery_fee = col_double(),
  last_pymnt_d = col_character(),
  last_pymnt_amnt = col_double(),
  next_pymnt_d = col_character(),
  last_credit_pull_d = col_character(),
  collections_12_mths_ex_med = col_double(),
  mths_since_last_major_derog = col_character(),
  policy_code = col_double(),
  application_type = col_character(),
  annual_inc_joint = col_character(),
  dti_joint = col_character(),
  verification_status_joint = col_character(),
  acc_now_delinq = col_double(),
  tot_coll_amt = col_character(),
  tot_cur_bal = col_character(),
  open_acc_6m = col_character(),
  open_il_6m = col_character(),
  open_il_12m = col_character(),
  open_il_24m = col_character(),
  mths_since_rcnt_il = col_character(),
  total_bal_il = col_character(),
  il_util = col_character(),
  open_rv_12m = col_character(),
  open_rv_24m = col_character(),
  max_bal_bc = col_character(),
  all_util = col_character(),
  total_rev_hi_lim = col_double(),
  inq_fi = col_character(),
  total_cu_tl = col_character(),
  inq_last_12m = col_character()
))
```

# EDA

```{r}
summary(data_lending)
```

Missing values

```{r}
missing_table <-  data_lending %>% 
  map_dbl(~ round(sum(is.na(.x))/length(.x), digits=2)) %>%
  data_frame(a=names(.), b=.) %>% 
  DT::datatable()

delete_missing <- missing_table$x$data %>%  filter(b>0.7) %>% select(a) %>% unlist()
data_lending_deleted_missing <- data_lending[ , -which(names(data_lending) %in% delete_missing)]

```
