---
title: "Lendingclub"
author: "elsa"
date: "8 septembre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
library(tidyverse)
library(data.table)
library(stringr)
library(DescTools)
library(lubridate)
```



```{r}
data_lending <- read_csv("../inputs/loan.csv", 
  col_types = cols(
  id = col_character(),
  member_id = col_character(),
  loan_amnt = col_double(),
  funded_amnt = col_double(),
  funded_amnt_inv = col_double(),
  term = col_character(),
  int_rate = col_double(),
  installment = col_double(),
  grade = col_character(),
  sub_grade = col_character(),
  emp_title = col_character(),
  emp_length = col_character(),
  home_ownership = col_character(),
  annual_inc = col_double(),
  verification_status = col_character(),
  issue_d = col_character(),
  loan_status = col_character(),
  pymnt_plan = col_character(),
  url = col_character(),
  desc = col_character(),
  purpose = col_character(),
  title = col_character(),
  zip_code = col_character(),
  addr_state = col_character(),
  dti = col_double(),
  delinq_2yrs = col_double(),
  earliest_cr_line = col_character(),
  inq_last_6mths = col_double(),
  mths_since_last_delinq = col_double(),
  mths_since_last_record = col_double(),
  open_acc = col_double(),
  pub_rec = col_double(),
  revol_bal = col_double(),
  revol_util = col_double(),
  total_acc = col_double(),
  initial_list_status = col_character(),
  out_prncp = col_double(),
  out_prncp_inv = col_double(),
  total_pymnt = col_double(),
  total_pymnt_inv = col_double(),
  total_rec_prncp = col_double(),
  total_rec_int = col_double(),
  total_rec_late_fee = col_double(),
  recoveries = col_double(),
  collection_recovery_fee = col_double(),
  last_pymnt_d = col_character(),
  last_pymnt_amnt = col_double(),
  next_pymnt_d = col_character(),
  last_credit_pull_d = col_character(),
  collections_12_mths_ex_med = col_double(),
  mths_since_last_major_derog = col_character(),
  policy_code = col_double(),
  application_type = col_character(),
  annual_inc_joint = col_character(),
  dti_joint = col_character(),
  verification_status_joint = col_character(),
  acc_now_delinq = col_double(),
  tot_coll_amt = col_character(),
  tot_cur_bal = col_character(),
  open_acc_6m = col_character(),
  open_il_6m = col_character(),
  open_il_12m = col_character(),
  open_il_24m = col_character(),
  mths_since_rcnt_il = col_character(),
  total_bal_il = col_character(),
  il_util = col_character(),
  open_rv_12m = col_character(),
  open_rv_24m = col_character(),
  max_bal_bc = col_character(),
  all_util = col_character(),
  total_rev_hi_lim = col_double(),
  inq_fi = col_character(),
  total_cu_tl = col_character(),
  inq_last_12m = col_character()
))
```

# Exploratory Data Analysis

```{r}
summary(data_lending)
```

## Missing values

```{r}
missing_table <-  data_lending %>% 
  map_dbl(~ round(sum(is.na(.x))/length(.x), digits=2)) %>%
  data_frame(a=names(.), b=.) %>% 
  DT::datatable()

delete_missing <- missing_table$x$data %>%  filter(b>0.7) %>% select(a) %>% unlist()

data_lending_deleted_missing <- data_lending[ , -which(names(data_lending) %in% delete_missing)]

missing_table

```

## Feature Cleaning

```{r}
#extract numbers from term
data_clean <- data_lending_deleted_missing %>% 
  mutate(term= as.numeric(str_extract_all
                          (term, "[[:digit:]]+")))

#extract numbers from employment length
data_clean1 <- data_clean %>% 
  mutate(emp_length= as.numeric(str_extract_all
                                (emp_length, "[[:digit:]]+")))

#Binary Verification Status
data_clean2 <- data_clean1 %>% 
  mutate(verification_status= ifelse
         (verification_status=="Not Verified", 0, 1)) 

#Issue Date into date format

months_dict <- c("01", "02", "03", "4", "5", "6", "7", "8", "9", "10", "11", "12")
names(months_dict) <- c("Jan", "Feb", "Mar", "Apr", "May",
                        "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

data_clean3 <- data_clean2 %>% 
  separate(issue_d, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(issue_d, d_d, good_m, y_d, sep="-") %>% 
  mutate(issue_d= dmy(issue_d)) %>% 
  select(-m_d)

#Earliest Credit Line into date format

data_clean4 <- data_clean3 %>% 
  separate(earliest_cr_line, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(earliest_cr_line, d_d, good_m, y_d, sep="-") %>% 
  mutate(earliest_cr_line= dmy(earliest_cr_line)) %>% 
  select(-m_d)

# Last payment into date format
  
data_clean5 <- data_clean4 %>% 
  separate(last_pymnt_d, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(last_pymnt_d, d_d, good_m, y_d, sep="-") %>% 
  mutate(last_pymnt_d= dmy(last_pymnt_d)) %>% 
  select(-m_d)

# Last Credit Pull into date format

data_clean6 <- data_clean5 %>% 
  separate(last_credit_pull_d, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(last_credit_pull_d, d_d, good_m, y_d, sep="-") %>% 
  mutate(last_credit_pull_d= dmy(last_credit_pull_d)) %>% 
  select(-m_d)
  
```


## Verification of mathematical formulas and Calculation of actual benefits for the investors

```{r}
# Formulas to calclulate monthly payment due 
data_clean6 %>% 
  mutate(monthly_pymt_calculated= 
           (loan_amnt*(int_rate*0.01/12))/(1 - (1 +int_rate*0.01/12) ^(-term))) %>%         mutate(diff_pmt=monthly_pymt_calculated - installment) %>% 
  select(installment, monthly_pymt_calculated, diff_pmt)

# Real total payment and benefits for the investors

data_benef <- data_clean6 %>%
  mutate(theo_pymt = installment*term) %>% 
  mutate(theo_benef = theo_pymt - loan_amnt)

# Investor theoretical and actual ROI

data_ROI <- data_benef %>% 
  mutate(theo_roi = (theo_benef/loan_amnt)) %>% 
  filter(loan_amnt==funded_amnt) %>% 
  mutate(diff_pymt = total_pymnt - total_pymnt_inv) %>% 
  mutate(real_roi= total_pymnt/loan_amnt)


```

# Graphs on loan amounts

```{r}
Desc(data_ROI$loan_amnt, main = "Loan amount distribution", plotit = TRUE)
```

## Total monthly amount in time 
```{r}
amnt_df <- data_ROI %>% 
  select(issue_d, loan_amnt) %>% 
  group_by(issue_d) %>% 
  summarise(Amount = sum(loan_amnt))

 ggplot(amnt_df,aes(x = issue_d, y = Amount)) +
  geom_line() + 
  labs(x="Date issued", y= "Loan Amount", title="Evolution of amount required in time")
```

## Proportion of amount 

## Amount compared to Grades 

```{r}
box_grade <- ggplot(data_ROI, aes(grade, loan_amnt))
box_grade + geom_boxplot(aes(fill = grade)) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Loan amount by grades",
    x = "Grades",
    y = "Amount")  
```


## Amount compared to status 

```{r}
box_status <- ggplot(data_ROI, aes(loan_status, loan_amnt))
box_status + geom_boxplot(aes(fill = loan_status)) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Loan amount by status",
    x = "Status",
    y = "Amount")

```



## Amount comparared to terms and grades

```{r}
box_term <- ggplot(data_ROI, aes(as.factor(term), loan_amnt))
box_term + geom_boxplot(aes(fill = as.factor(term))) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Loan amount by term and grades",
    x = "term",
    y = "Amount")

ggplot(data_ROI, aes(x= grade , y= loan_amnt, fill=as.factor(term) )) + 
  geom_bar(stat = "identity", position = "fill")


```


## Amount compared to purpose

```{r}
Desc(data_ROI$purpose, plotit = TRUE)

```


```{r}
box_purpose <- ggplot(data_ROI, aes(as.factor(purpose), loan_amnt))
box_purpose + geom_boxplot(aes(fill = as.factor(purpose))) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Loan amount by purpose",
    x = "term",
    y = "Amount")
```

## Loan amt compared to income

```{r}
data_inc <- data_ROI %>% 
  filter(annual_inc<500000) %>% 
  filter(loan_status!="Current") %>% 
  mutate(inc_amnt= loan_amnt/annual_inc)

ggplot(data_inc, aes(x=inc_amnt)) +
  geom_histogram(bins = 50)+
  facet_grid(.~ as.factor(term))+
  xlim(0, 0.7)

data_36 <- data_inc %>% 
  filter(term==36)
data_60 <- data_inc %>% 
  filter(term==60)
# 36 terms
ggplot(data_36, aes(x=annual_inc, y= loan_amnt)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 0.50, color="red")

#60 terms
ggplot(data_60, aes(x=annual_inc, y= loan_amnt)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 0.52, color="red")

ggplot(data_inc, aes(x=as.factor(term), y=annual_inc))+
  geom_boxplot()

```

## Est-ce que les gens qui empruntent beaucoup par rappor à leur salaire sont plus en défault et est-ce qu'ils ont les bonnes grades

```{r}
data_inc_factor <- data_inc %>% 
  mutate(term=as.factor(term)) %>% 
  mutate(loan_status= as.factor(loan_status))

ggplot(data_inc_factor, aes(x=loan_status, y=inc_amnt, fill=loan_status)) +
  geom_boxplot() +
  ylim(0,1) +
  facet_grid(.~term)

```

## ANOVA

```{r}
data_model_36 <- data_inc_factor %>% 
  filter(term==36) %>% 
  filter(loan_status %in% c("Fully Paid", "Charged Off"))

model_proportion <- lm(inc_amnt ~ loan_status, data=data_model_36)
summary(model_proportion)
anova(model_proportion)
confint(model_proportion)

data_model_60 <- data_inc_factor %>% 
  filter(term==60)


model_proportion <- lm(inc_amnt ~ loan_status, data=data_model_60)
summary(model_proportion)
anova(model_proportion)
confint(model_proportion)

```



```{r}
Desc(data_ROI$loan_status, plotit = TRUE)
```

## Theoretical Roi compared to Grades

```{r}
ggplot(data_ROI, aes(x= as.factor(loan_status), y=roi)) +
  geom_boxplot(aes(fill=as.factor(loan_status))) +
  theme(legend.position= "right")

```



## Theoretical Roi compared to actual status of loan

```{r}
ggplot(data_ROI, aes(x= as.factor(grade), y=roi)) +
  geom_boxplot(aes(fill=as.factor(grade))) +
  theme(legend.position= "right")

```

## Late payments 

```{r}
data_test <- data_ROI %>% 
  mutate(loan_end= issue_d + term*30.4375) %>% 
  filter(loan_status!="Current") %>% 
  mutate(late_days= loan_end - last_pymnt_d)

fully_paid <- data_test %>% 
  filter(loan_status=="Fully Paid")

ggplot(fully_paid, aes(x=round(as.numeric(late_days)/30.4375))) +
  geom_histogram(bins = max(fully_paid$late_days))

# plotter retard en fonciton de ROI
ggplot()

```


Quel proportion de mec voit son pret financer? 

Quel son les prêts qui font défaults

Rentabilité théorique vs réelle






